---
title: "Metabolomics: NAT16 F63S Variant and Histidine Metabolism in CKD"
author: "Naimisha Venkatesh"
date: "November 2025"
output:
  html_notebook:
    toc: true
    toc_float: true
    number_sections: true
---

**All data in this notebook are simulated.** 
# Setup

```{r install_packages, echo=TRUE}
#packages installation
required_packages <- c("ggplot2", "dplyr", "tidyr", "patchwork", "ggpubr", 
                       "pheatmap", "igraph", "ggraph", "reshape2")

# Check and install missing packages
missing_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]

if(length(missing_packages) > 0) {
  cat("Installing missing packages:", paste(missing_packages, collapse = ", "), "\n")
  install.packages(missing_packages, dependencies = TRUE)
  cat("Package installation complete\n\n")
} else {
  cat("All required packages are already installed\n\n")
}
invisible(lapply(required_packages, library, character.only = TRUE))

sample_data$Group <- factor(sample_data$Group, 
                            levels = c("Healthy_WT", "Healthy_Variant", 
                                      "CKD_WT", "CKD_Variant"))

group_colors <- c("Healthy_WT" = "#2E86AB", 
                  "Healthy_Variant" = "#A23B72",
                  "CKD_WT" = "#F18F01", 
                  "CKD_Variant" = "#C73E1D")


```

# Simulated data

## Sample table

Four groups of 30 samples each, crossing genotype (WT or F63S variant) with disease status (healthy or CKD).

```{r simulate_samples}

set.seed(123)
n_samples <- 120 #30 samples for each group

sample_data <- data.frame(SAMPLE_ID = paste0("S",1:n_samples), Group=rep(c("Healthy_WT","Healthy_Variant","CKD_WT","CKD_Variant"),each=30),Genotype = rep(c("WT", "Variant", "WT", "Variant"), each = 30),Disease = rep(c("Healthy", "Healthy", "CKD", "CKD"), each = 30))

#factor levels
sample_data$Group <- factor(sample_data$Group, 
                            levels = c("Healthy_WT", "Healthy_Variant", 
                                      "CKD_WT", "CKD_Variant"))
sample_data
```

## Gene expression

Simulated TPM-like expression values for NAT16 and five related genes: HAL, CARNS1, HDC, CNDP2, and NAT8L. Group means encode the expected effects, including reduced NAT16 in variant carriers and compensatory HAL upregulation in CKD.

```{r gene_expression, echo=TRUE}
genes <- c("NAT16", "HAL", "CARNS1", "HDC", "CNDP2", "NAT8L")
gene_expression <- data.frame(SAMPLE_ID = sample_data$SAMPLE_ID,
                              NAT16 = c(
    rnorm(30, mean = 50, sd = 8),   # Healthy WT
    rnorm(30, mean = 35, sd = 7),   # Healthy Variant (30% reduction)
    rnorm(30, mean = 45, sd = 9),   # CKD WT
    rnorm(30, mean = 25, sd = 6)    # CKD Variant (50% reduction)
  ),
  
  HAL = c(
    rnorm(30, mean = 30, sd = 5),
    rnorm(30, mean = 35, sd = 6),
    rnorm(30, mean = 40, sd = 7),
    rnorm(30, mean = 55, sd = 8)    # Compensatory upregulation
  ),
  
  CARNS1 = c(
    rnorm(30, mean = 40, sd = 6),
    rnorm(30, mean = 38, sd = 7),
    rnorm(30, mean = 35, sd = 6),
    rnorm(30, mean = 30, sd = 5)
  ),
  
  HDC = c(
    rnorm(30, mean = 25, sd = 4),
    rnorm(30, mean = 28, sd = 5),
    rnorm(30, mean = 32, sd = 6),
    rnorm(30, mean = 38, sd = 7)    # Upregulated in CKD
  ),
  
  CNDP2 = c(
    rnorm(30, mean = 35, sd = 5),
    rnorm(30, mean = 37, sd = 6),
    rnorm(30, mean = 42, sd = 7),
    rnorm(30, mean = 45, sd = 8)
  ),
  
  NAT8L = c(
    rnorm(30, mean = 45, sd = 7),
    rnorm(30, mean = 40, sd = 6),
    rnorm(30, mean = 38, sd = 7),
    rnorm(30, mean = 32, sd = 6)
  )
)
```

## Metabolites

N-acetylhistidine and histidine are drawn directly from group-specific distributions. The remaining metabolites are calculated from the simulated gene expression plus noise, so that downstream gene-metabolite correlations are structured rather than random.

```{r metabolite_data, echo=TRUE}
metabolite_data <- data.frame(
  SAMPLE_ID = sample_data$SAMPLE_ID,
  
  # Primary metabolites (targeted metabolomics)
  N_Acetylhistidine = c(
    rnorm(30, mean = 15.2, sd = 2.3),  # Healthy WT
    rnorm(30, mean = 9.8, sd = 1.8),   # Healthy Variant (35% reduction)
    rnorm(30, mean = 8.5, sd = 2.1),   # CKD WT (44% reduction)
    rnorm(30, mean = 4.2, sd = 1.5)    # CKD Variant (72% reduction)
  ),
  
  Histidine = c(
    rnorm(30, mean = 78, sd = 12),     # Healthy WT
    rnorm(30, mean = 82, sd = 15),     # Healthy Variant
    rnorm(30, mean = 95, sd = 18),     # CKD WT (elevated)
    rnorm(30, mean = 118, sd = 22)     # CKD Variant (highly elevated)
  ),
  
  # Link metabolites to gene expression for realistic correlations
  Carnosine = gene_expression$CARNS1 * 0.4 - (gene_expression$CNDP2 * 0.2) + rnorm(n_samples, 0, 3),
  Histamine = gene_expression$HDC * 0.25 + rnorm(n_samples, 0, 1.5),
  Urocanate = gene_expression$HAL * 0.35 + rnorm(n_samples, 0, 2),
  
  Methylhistidine_1 = 25 + (120 - gene_expression$NAT16) * 0.15 + rnorm(n_samples, 0, 2),
  Methylhistidine_3 = 28 + (120 - gene_expression$NAT16) * 0.18 + rnorm(n_samples, 0, 2.5),
  
  # Oxidative stress and inflammation markers
  Oxidative_Stress = 50 - (gene_expression$NAT16 * 0.3) - (gene_expression$CARNS1 * 0.2) + rnorm(n_samples, 0, 4),
  PGE2 = 15 + (sample_data$Disease == "CKD") * 8 + rnorm(n_samples, 0, 2),
  MDA = 12 + (sample_data$Disease == "CKD") * 6 + rnorm(n_samples, 0, 1.5)
)

# Calculate ratio
metabolite_data$Ratio <- metabolite_data$Histidine / metabolite_data$N_Acetylhistidine

# Generate eGFR (kidney function)
metabolite_data$eGFR <- c(
  rnorm(30, mean = 95, sd = 10),   # Healthy WT
  rnorm(30, mean = 92, sd = 12),   # Healthy Variant
  rnorm(30, mean = 35, sd = 8),    # CKD WT
  rnorm(30, mean = 28, sd = 7)     # CKD Variant
)

# Merge all datasets
integrated_data <- sample_data %>%
  left_join(gene_expression, by = "SAMPLE_ID") %>%
  left_join(metabolite_data, by = "SAMPLE_ID")

# Print dataset
metabolite_data
```

# Figures

## Figure 1: Metabolite levels by group

Boxplots showing which metabolites increase or decrease in each group, with t-test comparisons between healthy-WT and CKD-variant. Panel D relates N-acetylhistidine to kidney function.

```{r figure1_boxplots, echo=TRUE, warning=FALSE, message=FALSE}
# Panel A: N-acetylhistidine levels
plot_A <- ggplot(integrated_data, aes(x = Group, y = N_Acetylhistidine, fill = Group)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.5) +
  scale_fill_manual(values = group_colors) +
  labs(title = "A. N-Acetylhistidine Levels",
       y = "N-Acetylhistidine (μM)",
       x = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  stat_compare_means(comparisons = list(c("Healthy_WT", "CKD_Variant")),
                     label = "p.signif", method = "t.test")

# Panel B: Histidine levels
plot_B <- ggplot(integrated_data, aes(x = Group, y = Histidine, fill = Group)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.5) +
  scale_fill_manual(values = group_colors) +
  labs(title = "B. Histidine Levels",
       y = "Histidine (μM)",
       x = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  stat_compare_means(comparisons = list(c("Healthy_WT", "CKD_Variant")),
                     label = "p.signif", method = "t.test")

# Panel C: Ratio
plot_C <- ggplot(integrated_data, aes(x = Group, y = Ratio, fill = Group)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.5) +
  scale_fill_manual(values = group_colors) +
  labs(title = "C. Histidine/N-Acetylhistidine Ratio",
       y = "His/N-AcHis Ratio",
       x = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  stat_compare_means(comparisons = list(c("Healthy_WT", "CKD_Variant")),
                     label = "p.signif", method = "t.test")

# Panel D: Correlation with eGFR
plot_D <- ggplot(integrated_data, aes(x = eGFR, y = N_Acetylhistidine, color = Group)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = group_colors) +
  labs(title = "D. Correlation with Kidney Function",
       x = "eGFR (mL/min/1.73m²)",
       y = "N-Acetylhistidine (μM)") +
  theme(legend.position = "bottom",
        legend.title = element_blank())

# Combine all panels
figure_1 <- (plot_A | plot_B) / (plot_C | plot_D) +
  plot_annotation(
    title = "NAT16 Variant Effects on Histidine Metabolism",
    subtitle = "Metabolite levels across health status and genotype",
    theme = theme(plot.title = element_text(size = 14, face = "bold"))
  )

print(figure_1)

```

## Figure 2A: Gene-metabolite correlation heatmap

Pearson correlations between the six genes and eight metabolites, with per-pair `cor.test` p-values shown as significance stars and both axes hierarchically clustered. Pairs exceeding |r| > 0.5 at p < 0.05 are tabulated above the heatmap.

```{r figure2a_heatmap, echo=TRUE, warning=FALSE, message=FALSE}
gene_matrix <- integrated_data %>% select(all_of(genes))
metabolite_matrix <- integrated_data %>% 
  select(N_Acetylhistidine, Histidine, Carnosine, Histamine, Urocanate,
         Methylhistidine_1, Methylhistidine_3, Oxidative_Stress)

# Calculate correlation matrix
gene_metabolite_cor <- cor(gene_matrix, metabolite_matrix, method = "pearson")

# Calculate p-values
cor_pvalues <- matrix(NA, nrow = nrow(gene_metabolite_cor), ncol = ncol(gene_metabolite_cor))
rownames(cor_pvalues) <- rownames(gene_metabolite_cor)
colnames(cor_pvalues) <- colnames(gene_metabolite_cor)

for(i in 1:nrow(gene_metabolite_cor)) {
  for(j in 1:ncol(gene_metabolite_cor)) {
    test <- cor.test(gene_matrix[,i], metabolite_matrix[,j])
    cor_pvalues[i,j] <- test$p.value
  }
}

# Identify significant correlations
significant_cors <- which(abs(gene_metabolite_cor) > 0.5 & cor_pvalues < 0.05, arr.ind = TRUE)
sig_cor_df <- data.frame(
  Gene = rownames(gene_metabolite_cor)[significant_cors[,1]],
  Metabolite = colnames(gene_metabolite_cor)[significant_cors[,2]],
  Correlation = gene_metabolite_cor[significant_cors],
  P_value = cor_pvalues[significant_cors]
)

cat("\n=== Significant Gene-Metabolite Correlations (|r| > 0.5, p < 0.05) ===\n")
print(sig_cor_df %>% arrange(desc(abs(Correlation))))
cat("\n")

# Create significance annotation matrix
sig_matrix <- ifelse(cor_pvalues < 0.001, "***",
                     ifelse(cor_pvalues < 0.01, "**",
                            ifelse(cor_pvalues < 0.05, "*", "")))

# Clean column names for display
colnames(gene_metabolite_cor) <- gsub("_", "-", colnames(gene_metabolite_cor))
colnames(sig_matrix) <- gsub("_", "-", colnames(sig_matrix))

pheatmap(gene_metabolite_cor,
         display_numbers = sig_matrix,
         number_color = "black",
         color = colorRampPalette(c("#2E86AB", "white", "#C73E1D"))(100),
         breaks = seq(-1, 1, length.out = 101),
         main = "Visualization 2A: Gene-Metabolite Correlation Heatmap",
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         fontsize = 11,
         fontsize_number = 10,
         angle_col = 45,
         border_color = "grey60")


```

## Figure 2A: Volcano plot

A volcano plot of 150 metabolites contrasting CKD-variant against healthy-WT. Ten pathway-relevant metabolites are given fold changes and p-values reflecting the expected direction of change; the remaining 140 are background, generated randomly to give the plot realistic density.

```{r volcano_plot, warning=FALSE, message=FALSE}
set.seed(123)
n_metabolites <- 150

metabolite_names <- c(
  "N-Acetylhistidine", "Carnosine", "Histidine", "1-Methylhistidine",
  "3-Methylhistidine", "Histamine", "8-OHdG", "MDA", "PGE2", "PGF2α",
  paste0("Metabolite_", 1:(n_metabolites - 10))
)

# Generate fold changes (log2)
log2FC <- rnorm(n_metabolites, mean = 0, sd = 0.8)
log2FC[1] <- log2(1/3.2)
log2FC[2] <- log2(1/1.8)
log2FC[3] <- log2(1.5)
log2FC[4] <- log2(1.7)
log2FC[5] <- log2(1.7)
log2FC[7] <- log2(2.1)
log2FC[8] <- log2(2.1)
log2FC[9] <- log2(1.9)
log2FC[10] <- log2(1.9)

# Generate p-values
p_values <- runif(n_metabolites, 0.001, 0.5)
p_values[1] <- 0.0001
p_values[2] <- 0.02
p_values[3] <- 0.001
p_values[4] <- 0.005
p_values[5] <- 0.005
p_values[7] <- 0.0005
p_values[8] <- 0.0005
p_values[9] <- 0.001
p_values[10] <- 0.001

volcano_data <- data.frame(
  Metabolite = metabolite_names,
  log2FC = log2FC,
  pvalue = p_values,
  neg_log10p = -log10(p_values)
)

volcano_data <- volcano_data %>%
  mutate(
    Category = case_when(
      Metabolite %in% c("N-Acetylhistidine", "Carnosine", "Histidine", 
                        "1-Methylhistidine", "3-Methylhistidine") ~ "NAT16-related",
      Metabolite %in% c("8-OHdG", "MDA") ~ "Oxidative stress",
      Metabolite %in% c("PGE2", "PGF2α") ~ "Inflammation",
      abs(log2FC) > 1 & pvalue < 0.05 ~ "Significant",
      TRUE ~ "Not significant"
    ),
    Label = ifelse(
      Metabolite %in% c("N-Acetylhistidine", "Carnosine", "Histidine", 
                        "1-Methylhistidine", "8-OHdG", "PGE2"),
      Metabolite, ""
    )
  )

volcano_plot <- ggplot(volcano_data, aes(x = log2FC, y = neg_log10p)) +
  geom_point(aes(color = Category, size = Category), alpha = 0.6) +
  scale_color_manual(
    values = c(
      "NAT16-related" = "#C73E1D",
      "Oxidative stress" = "#F18F01",
      "Inflammation" = "#A23B72",
      "Significant" = "#2E86AB",
      "Not significant" = "gray70"
    )
  ) +
  scale_size_manual(
    values = c(
      "NAT16-related" = 3,
      "Oxidative stress" = 2.5,
      "Inflammation" = 2.5,
      "Significant" = 2,
      "Not significant" = 1.5
    )
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "gray40") +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "gray40") +
  geom_text(aes(label = Label), hjust = -0.1, vjust = 0, size = 3, 
            check_overlap = TRUE) +
  labs(
    title = "Metabolomic Signature of NAT16 F63S Variant in CKD",
    subtitle = "CKD-Variant vs Healthy-WT",
    x = "log2(Fold Change)",
    y = "-log10(p-value)"
  ) +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  ) +
  annotate("text", x = -2, y = 1, label = "Decreased", color = "gray40", size = 4) +
  annotate("text", x = 2, y = 1, label = "Increased", color = "gray40", size = 4)

print(volcano_plot)


```

## Figure 2B: Metabolic pathway diagram

Fold change values are set the same way as in the volcano plot above. The diagram gives a visual map of how the NAT16 F63S variant is expected to disrupt histidine metabolism and trigger a downstream cascade.

The decrease in N-acetylhistidine (blue, center) represents the primary enzymatic deficiency, while histidine accumulation (red, left) reflects substrate-product imbalance. Loss of N-acetylhistidine is proposed to lead to elevated oxidative stress (2.1-fold) and inflammation (1.9-fold), offering a candidate mechanism for increased CKD susceptibility in variant carriers.

```{r pathway_diagram, warning=FALSE, message=FALSE}

pathway_data <- data.frame(
  Metabolite = c("Histidine", "N-Acetylhistidine", "Histamine", 
                 "Carnosine", "Oxidative\nStress", "Inflammation"),
  x = c(1, 3, 1.5, 1.5, 4, 4),
  y = c(3, 3, 1.5, 4.5, 2, 4),
  FoldChange = c(1.5, -3.2, 1.2, -1.8, 2.1, 1.9)
)

connections <- data.frame(
  x = c(1, 1, 1, 3, 3),
  y = c(3, 3, 3, 3, 3),
  xend = c(3, 1.5, 1.5, 4, 4),
  yend = c(3, 1.5, 4.5, 2, 4)
)

pathway_plot <- ggplot() +
  geom_segment(data = connections, 
               aes(x = x, y = y, xend = xend, yend = yend),
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
               linewidth = 1, color = "gray40") +
  annotate("text", x = 2, y = 3.3, label = "NAT16\n(F63S variant)", 
           size = 3.5, fontface = "bold", color = "#C73E1D") +
  geom_point(data = pathway_data, 
             aes(x = x, y = y, fill = FoldChange, size = abs(FoldChange)),
             shape = 21, color = "black", stroke = 1.5) +
  scale_fill_gradient2(
    low = "#2E86AB", mid = "white", high = "#C73E1D",
    midpoint = 0, limits = c(-3.5, 2.5),
    name = "Fold Change"
  ) +
  scale_size_continuous(range = c(8, 15), guide = "none") +
  geom_text(data = pathway_data, 
            aes(x = x, y = y, label = Metabolite),
            size = 3, fontface = "bold") +
  labs(
    title = "Integrated Metabolic Pathway Analysis",
    subtitle = "NAT16 dysfunction triggers cascade of metabolic changes"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 10, hjust = 0.5),
    legend.position = "right"
  ) +
  coord_fixed(ratio = 1) +
  xlim(0.5, 4.5) +
  ylim(0.5, 5)

print(pathway_plot)


```

# Alternative visualizations

## Figure 3: Scatter plots by disease status

The NAT16-metabolite relationship shown separately for healthy and CKD samples, faceted by metabolite and colored by genotype.

```{r figure3_facet_scatter, warning=FALSE, message=FALSE}
# Select key metabolites to visualize
key_metabolites <- c("N_Acetylhistidine", "Histidine", "Carnosine", 
                     "Oxidative_Stress", "Histamine", "Urocanate")

# Reshape data for faceting
plot_data <- integrated_data %>%
  select(SAMPLE_ID, Group, Disease, Genotype, NAT16, all_of(key_metabolites)) %>%
  pivot_longer(cols = all_of(key_metabolites), 
               names_to = "Metabolite", 
               values_to = "Concentration")

# Create faceted scatter plot
facet_scatter <- ggplot(plot_data, aes(x = NAT16, y = Concentration, color = Genotype)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1) +
  facet_grid(Metabolite ~ Disease, scales = "free_y") +
  scale_color_manual(values = c("WT" = "#2E86AB", "Variant" = "#C73E1D")) +
  labs(
    title = "NAT16 Expression vs Metabolite Levels by Disease Status",
    subtitle = "Shows how NAT16-metabolite relationships differ in CKD",
    x = "NAT16 Expression (TPM)",
    y = "Metabolite Concentration (μM)"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom"
  )

print(facet_scatter)

# Save
ggsave("Figure3_FacetedScatter.png", facet_scatter, width = 12, height = 14, dpi = 300)
```

## Figure 4: Correlation coefficient plot

Correlation strength for each NAT16-metabolite pair, computed separately in healthy and CKD samples to show how the relationships shift with disease.

```{r figure4_cor_dotplot, warning=FALSE, message=FALSE}

# Calculate correlations separately for Healthy and CKD
calc_cors <- function(data, disease_status) {
  subset_data <- data %>% filter(Disease == disease_status)
  
  gene_mat <- subset_data %>% select(all_of(genes))
  met_mat <- subset_data %>% select(N_Acetylhistidine, Histidine, Carnosine, 
                                    Histamine, Urocanate, Oxidative_Stress)
  
  cors <- cor(gene_mat, met_mat, method = "pearson")
  
  # Convert to long format
  cors_long <- as.data.frame(cors) %>%
    rownames_to_column("Gene") %>%
    pivot_longer(-Gene, names_to = "Metabolite", values_to = "Correlation") %>%
    mutate(Disease = disease_status)
  
  return(cors_long)
}

# Calculate for both groups
library(tibble)
healthy_cors <- calc_cors(integrated_data, "Healthy")
ckd_cors <- calc_cors(integrated_data, "CKD")
all_cors <- bind_rows(healthy_cors, ckd_cors)

# Focus on NAT16 only for cleaner visualization
nat16_cors <- all_cors %>% filter(Gene == "NAT16")

# Create dot plot
cor_dotplot <- ggplot(nat16_cors, aes(x = Correlation, y = Metabolite, color = Disease)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(size = 4, alpha = 0.8) +
  geom_line(aes(group = Metabolite), color = "gray70", linewidth = 0.5) +
  scale_color_manual(values = c("Healthy" = "#2E86AB", "CKD" = "#C73E1D")) +
  scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, 0.25)) +
  labs(
    title = "NAT16-Metabolite Correlations: Healthy vs CKD",
    subtitle = "Shows how correlations change in disease state",
    x = "Pearson Correlation Coefficient",
    y = NULL
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_line(color = "gray90"),
    axis.text.y = element_text(size = 11, face = "bold")
  )

print(cor_dotplot)

# Save
ggsave("Figure4_CorrelationDotPlot.png", cor_dotplot, width = 10, height = 6, dpi = 300)
```

# Summary statistics and export

Per-group means and standard deviations for the key metabolites and eGFR, a global correlation test between eGFR and N-acetylhistidine, and export of the merged data, correlation tables, and figures.

```{r summary_stats, warning=FALSE, message=FALSE}
summary_stats <- integrated_data %>%
  group_by(Group) %>%
  summarise(
    N_AcHis_mean = mean(N_Acetylhistidine),
    N_AcHis_sd = sd(N_Acetylhistidine),
    His_mean = mean(Histidine),
    His_sd = sd(Histidine),
    Ratio_mean = mean(Ratio),
    Ratio_sd = sd(Ratio),
    eGFR_mean = mean(eGFR),
    eGFR_sd = sd(eGFR)
  )

print(summary_stats)

# Correlation analysis
cor_test <- cor.test(integrated_data$eGFR, integrated_data$N_Acetylhistidine)
cat("\nCorrelation between eGFR and N-Acetylhistidine:\n")
cat(paste("r =", round(cor_test$estimate, 3), ", p-value =", 
          format.pval(cor_test$p.value, digits = 3), "\n"))

# Export data
write.csv(gene_metabolite_cor, "gene_metabolite_correlations.csv", row.names = TRUE)
write.csv(sig_cor_df, "significant_correlations.csv", row.names = FALSE)
write.csv(integrated_data, "integrated_multiomics_data.csv", row.names = FALSE)
# Save plots
 ggsave("Figure1_BoxPlots.png", figure_1, width = 12, height = 10, dpi = 300)
 ggsave("Figure2A_VolcanoPlot.png", volcano_plot, width = 10, height = 8, dpi = 300)
 ggsave("Figure2B_PathwayDiagram.png", pathway_plot, width = 10, height = 8, dpi = 300)


```
